---
title: "data_manipulation"
output: github_document
date: "2022-09-22"
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyverse)
library(readr)

```

## Import Dataset

```{r}
litters_df = read_csv("data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)

pups_data = read_csv("data/FAS_pups.csv")
pups_data = janitor::clean_names(pups_data)
```

## Select

```{r, echo=FALSE}
select(litters_df, group, litter_number, gd0_weight, pups_born_alive)
#select(desired ddf, 4 columns you want)
select(litters_df, group:gd_of_birth)
# this is the to select between group and birth

# you can also remove which variables you want to remove
select(litters_df, -pups_survive, -group)
# this means keep everything but survive and group. 

```


# Rename Variables when select
```{r}
select(litters_df, GROUP = group, LiTtEr_NuMbEr = litter_number)
#rename as part of the selection process

rename(litters_df, GROUP = group, LiTtEr_NuMbEr = litter_number)
#rename without selecting
```

# select with functions
```{r}
select(litters_df, starts_with("gd"))
select(litters_df, ends_with("ve"))

select(litters_df, litter_number, pups_survive, everything())
#this function put litter number and survive up at the front, and put everything else after. This function changes the order of the variables. 

relocate(litters_df, litter_number, pups_survive)
#This relocate function does the same reordering as the above. 
```


## `filter` select certain rows/ certain observatons.

some examples:
`gd_of_birth == 20`
`pups_born_alive >= 2`
`pups_survive != 4`
`!(pups_survive == 4)`
`group %in% c("Con7", "Con8")`
`group == "Con7" & gd_of_birth == 20`
`!((pups_survive == 4) & (gd_of_birth == 20))`

```{r}
filter(litters_df, group == "Con7" & gd_of_birth == 20)
```

`drop_na(litters_df)` will remove any row with a missing value
`drop_na(litters_df, wt_increase)` will remove rows for which wt_increase is missing.

```{r}
drop_na(litters_df)
drop_na(litters_df, gd18_weight)
```

## `mutate`
used when you want to create new variables, change current variabels. 

```{r}
mutate(litters_df,
  wt_gain = gd18_weight - gd0_weight,
  wt_gain_kg = wt_gain * 2.2, 
  group = str_to_lower(group)
)

#str_to_lower: is changing the variable to lower case. 
```


##`arrange`
rearrange the rows in your data. 
eg. sort your rows by weight, then it goes from lower weight to higher weight. 
```{r}
head(arrange(litters_df, group, pups_born_alive), 10)
```


## Piping '%>%' 

to do piping, we can create several intermediate datasets untill you get to the final dataset. 
```{r}
litters_data_raw = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data_clean_names = janitor::clean_names(litters_data_raw)
litters_data_selected_cols = select(litters_data_clean_names, -pups_survive)
litters_data_with_vars = 
  mutate(
    litters_data_selected_cols, 
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group))
litters_data_with_vars_without_missing = 
  drop_na(litters_data_with_vars, wt_gain)
litters_data_with_vars_without_missing
```

or, we could nest our functions:
In this case, the first step is in the middle of the code, then it reads from the middle to the outside. 
```{r}
litters_data_clean = 
  drop_na(
    mutate(
      select(
        janitor::clean_names(
          read_csv("./data/FAS_litters.csv", col_types = "ccddiiii")
          ), 
      -pups_survive
      ),
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)
    ),
  wt_gain
  )

litters_data_clean
```

Instead, we are using the pipe operator `%>%`
shortcut: "command + shift + m"
```{r}
litters_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)) %>% 
  drop_na(wt_gain)

litters_data 
```

