---
title: "Visualization with ggplot2_lecture9.29"
author: "Fengdi Zhang"
date: "2022-09-29"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
```

## Import Data 

```{r cars}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## Basic Scatterplot

```{r pressure, echo=FALSE}
ggplot(weather_df, aes(x = tmin, y = tmax)) + geom_point()
```

The same scatterplot with piping code:
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```
This is a good way to do when you want to modify the data before plotting. 

You can also save the output of ggplot() to an object and modify / print it later
```{r}
plot_weather = 
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) 

plot_weather + geom_point()
```


##Advanced scatterplot

adding color
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  geom_smooth(se = FALSE)
```
`aes` astheric mapping
`color = name` differnt color by name
`geom_smooth` adding a line to describe the trend. 
`alpha = .5` defines how transparent the dots are. 

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point() +
  geom_smooth(se = FALSE)
```
compare this to the above code, the `color = name` position is different, returning diff graphs. 



If you want to make seperate panels for each location by `facet_grid`;
```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```
Within `facet`, you use `~` to define the vairable you want on both side. 
Try `.~name` and `name~.`, then you would see the difference. 


## Some quick stuff
```{r}
ggplot(weather_df, aes(x = tmax, y = tmin)) + 
  geom_hex()
```
This shows how many points are in one grid. 


