---
title: "lecture 10.4_ vidualization2"
author: "Fengdi Zhang"
date: "2022-10-04"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)
library(patchwork)
```

## 


```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## adding labels, colors

```{r}
ggp_temp_plot = 
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15ยบ C", "0", "15")) +
    viridis::scale_color_viridis(
      name = "Location",
      discrete = TRUE
    )
```
`viridis::scale_color_viridis` viridis is a color scale, there are many diff colour scales can search online. `name = "Location"` tells what variable you want to use diff color. 

##Themes

```{r}
ggp_temp_plot + 
  theme_minimal() + 
  theme(legend.position = "bottom")
```
put the theme at the bottom allows greater left to right position. 
`theme_bw` `theme_minimal` `theme_classic` can change the background grid. 
the order of the `theme_minimal` and `legend` matters. 

## Data in geom()

```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park)
```
can use two diff datasets in one graph by adding in `geom` seperately

## Patchwork

Sometimes you want to put multiple graphs together,eg. in publication

```{r}
tmax_tmin_p = 
  weather_df %>% 
  ggplot(aes(x = tmax, y = tmin, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

(tmax_tmin_p + prcp_dens_p) / tmax_date_p
```
this function is under the packet `patchwork`. 
`A + B` means A and B side by side
`A / B` means A on the top of B

## Data Manipulation

R is seeing the factor and character variables as numbers, such as 1, 2, 3. 
so by default, 

```{r}
weather_df %>%
  mutate(name = forcats::fct_relevel(name, c("Waikiki_HA", "CentralPark_NY", "Waterhole_WA"))) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```
The `mutate` function is reordering the location names as desired. 

```{r}
weather_df %>%
  mutate(name = fct_reorder(name, tmax)) %>% 
  ggplot(aes(x = name, y = tmax))+
  geom_boxplot()
```
This `mutate` is to recorder location names by the median of `tmax`. 


```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  select(id, visit, everything()) %>%
  mutate(
    visit = recode(visit, "bl" = "00m"),
    visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) %>%
  arrange(id, visit)

ggplot(pulse_data, aes(x = visit, y = bdi)) + 
  geom_boxplot()
```
`pivot_longer` rearrange the dataset structure
