---
title: "Lecture10.6_numeric EDA"
author: "Fengdi Zhang"
date: "2022-10-6"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

## import data

```{r}
weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())
```
`lubridate` packege help with date. 
`floor_date` round date down. 2.27 --> 2.1; 3.8 --> 3.1, etc
we are doing this because we will `groupby` month later today. 

## Group_by()

we will group by the name first, 
```{r}
weather_df %>%
  group_by(name, month)
```
The output shows that there are 3*12 groups now. Although nothing has beed changed to the dataset. 

## Summarize ()
Let's group and then summarize

```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(n_obs = n())
```
`n_obs = n()` naming a new variable `n_obs` using the function `n()` to count how many subjects there are inside each group. 

count in other ways, not doing groupby and summarize
```{r}
weather_df %>%
  count(name, month)
```

don't use `table()` because it is not forming a dataframe, it doesn't allow you to do any analysis. 
```{r, eval = FALSE}
weather_df %>%
  pull(month) %>% 
  table()
```

You can have multiple summaries
```{r}
weather_df %>%
  group_by(month) %>%
  summarize(
    n_obs = n(),
    n_days = n_distinct(date))
```
`n_distinct()` counts how many distinct subjects there are in the variable. 

## A brief aside on 2x2 tables

```{r}
weather_df %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""  #tells whatever didn't meet the last two lines
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  group_by(name, cold) %>% 
  summarize(count = n()) %>% 
  pivot_wider(
    names_from = cold,
    values_from = count
  )
```


```{r}
weather_df %>% 
  mutate(cold = case_when(
    tmax <  5 ~ "cold",
    tmax >= 5 ~ "not_cold",
    TRUE     ~ ""
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  janitor::tabyl(name, cold)
```
`tabyl` function in `janitor` package gives a quicker way to create 2x2 table. 

`janitor` package is very helpful to do lots of things. count total rows, columns, tests, etc. 


## General Summaries
```{r}
weather_df %>%
  group_by(month) %>%
  summarize(
    mean_tmax = mean(tmax), #computing the mean of tmax in every group
    mean_prec = mean(prcp, na.rm = TRUE), 
    # na.rm = TRUE means we are removing the NAs/missing values. 
    median_tmax = median(tmax),
    sd_tmax = sd(tmax, na.rm = TRUE))
```

If you want to summarize multiple columns using the same summary, the across function is helpful.
```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(across(tmin:prcp, mean))
  # you are computing the mean for each variable between tmin and prcp. 
```

```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax, na.rm = TRUE)) %>% 
  #a dataframe has been created here, then you can start plottin. 
  ggplot(aes(x = month, y = mean_tmax, color = name)) + 
    geom_point() + geom_line() + 
    theme(legend.position = "bottom")
```


```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax)) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax) %>% 
  knitr::kable(digits = 1) #this tells R to form a table when knitting
```

##Grouped mutates

```{r}
weather_df %>%
  group_by(name) %>%
  mutate( #here the mutate function is conducted for each group
    mean_tmax = mean(tmax, na.rm = TRUE),
    centered_tmax = tmax - mean_tmax) %>% 
  ggplot(aes(x = date, y = centered_tmax, color = name)) + 
    geom_point() 
```


## Window functions 

Lagged observations
```{r}
weather_df %>%
  group_by(name) %>%
  mutate(
    yesterday_tmax = lag(tmax),
    tmax_change = tmax - yesterday_tmax) %>% #this could show the day to day change of temp.
    summarize(
      sd_tmax_change = sd(tmax_change, na.rm = TRUE)
    )
```

`lead` is going to the opposite direction of `lag`. 
```{r}
weather_df %>%
  group_by(name) %>%
  mutate(
    yesterday_tmax = lead(tmax))
```


Another function  `rank`
```{r}
weather_df %>%
  group_by(name, month) %>%
  mutate(temp_ranking = min_rank(tmax))
```

We could, for example, keep only the day with the lowest max temperature within each month:
```{r}
weather_df %>%
  group_by(name, month) %>%
  mutate(temp_ranking = min_rank(tmax)) %>% 
  filter(min_rank(tmax) < 2) %>% 
  arrange(name, month, temp_ranking)
```


