---
title: "Lecture11.10_LinearModel"
author: "Fengdi Zhang"
date: "2022-11-14"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

set.seed(1)

data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb %>% 
  mutate(stars = review_scores_location / 2) %>% 
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood) %>% 
  filter(borough != "Staten Island") %>% 
  select(price, stars, borough, neighborhood, room_type)
```

## Model Fitting

```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)

summary(fit)
```
`lm` means linear model
`price ~ stars + borough` means price is the outcome, stars and boroughs are predictors. 

## Get clean outputs from a linear model
Use `broom::tidy`
```{r}
fit %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  mutate(term = str_replace(term, "^borough", "Borough: ")) %>% 
  knitr::kable(digits = 3)

fit %>% 
  broom::glance()
#take a glance
```

Let's change reference category to Manhattan
```{r}
fit = nyc_airbnb %>% 
  mutate(
    borough = fct_infreq(borough) 
    #order the categories first when cleaning data, it always takes the first group
    #as the reference group
  ) %>% 
  lm(price ~ stars + borough, data = .)

fit %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  mutate(term = str_replace(term, "^borough", "Borough: ")) %>% 
  knitr::kable(digits = 3)
```

## Diagnostics
`modelr::add_residuals` is a function to add residuals in a dataset. 
```{r}
modelr::add_residuals(nyc_airbnb, fit) 

##Then you can plot some graphs to look at the residules. 
nyc_airbnb %>% 
  modelr::add_residuals(fit) %>% 
  ggplot(aes(x = borough, y = resid)) + geom_violin() + ylim(-250,250)

nyc_airbnb %>% 
  modelr::add_residuals(fit) %>% 
  ggplot(aes(x = stars, y = resid)) + geom_point()
```


## Hypothesis Testing

One coefficient(Let's say  `stars`)

When you are not sure if you need to include room_type as a predictor, Then you need to run ANOVA test. 

First, let's fit the null and alternative models. 
```{r}
fit_null = lm(price ~ stars + borough, data = nyc_airbnb)
fit_alt = lm(price ~ stars + borough + room_type, data = nyc_airbnb)
```
Then run ANOVA
```{r}
anova(fit_null, fit_alt) %>% 
  broom::tidy()
## The P-value is very small, so you need to inlcude room_type in the model. 
```


## Room_type by borough

Interactions...?

```{r}
nyc_airbnb %>% 
  lm(price ~ stars + room_type * borough, data = .) %>% 
  broom::tidy() %>% 
  knitr::kable(digits = 3)
```

so... can we fit models by borough ...?

```{r}
nest_lm_res =
  nyc_airbnb %>% 
  nest(df = -borough) %>% 
  mutate(
    models = map(.x = df, ~lm(price ~ stars + room_type, data = .x)),
    results = map(models, broom::tidy)) %>% 
  select(-df, -models) %>% 
  unnest(results)

# Now, I have borough specific model estimates. 

# The results of this approach are given in the table below.
nest_lm_res %>% 
  select(borough, term, estimate) %>% 
  mutate(term = fct_inorder(term)) %>% 
  pivot_wider(
    names_from = term, values_from = estimate) %>% 
  knitr::kable(digits = 3)
```

Quick double check ...?

```{r}

```





