---
title: "lecture 10.18_String and Factor Variable"
author: "Fengdi Zhang"
date: "2022-10-18"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
library(p8105.datasets)
```

##    String Vectors

```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "jeff")
str_detect(string_vec, "a")
str_detect(string_vec, "m")

str_replace(string_vec, "jeff", "Jeff")
```
everything for string start with `str`
`str_detect` is detecting anything that contains the element you are looking for
`str_replace` can replace what you want. 


```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "^i think")
str_detect(string_vec, "i think$")
```
`^` means find exactly what begins with `i think`. 
`$` means find exactly what ends with `i think`. 

```{r}
string_vec = c(
  "Y'all remember Pres. HW Bush?",
  "I saw a green bush",
  "BBQ and Bushwalking at Molonglo Gorge",
  "BUSH -- LIVE IN CONCERT!!"
  )

str_detect(string_vec,"[Bb]ush")
```
`[Bb]` means to detect anything that start with either B or b. 

```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "^[0-9][a-zA-Z]")
```
`^[0-9][a-zA-Z]` means to detect the ones that start with any number between 0-9 and any letter. 

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")
```
The character `.` matches anything, but there has to be something, there cannot be nothing. 

```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[[0-9]")
```

If you want to search for the actual bracket `[` or anctual dot `.`?
Then you need to use 2 backslash lines in front, such as `\\[` means to search for the actual bracket. 
`\\[[0-9]` means search for a bracket followed by a number. 

## Why factors are weird
```{r}
vec_sex = factor(c("male", "male", "female", "female"))
vec_sex
```
Once the string is converted into a factor, then the levels will have order by alphabetic order. 

```{r}
vec_sex = fct_relevel(vec_sex, "male")
vec_sex
```
This is to relevel the sex to start with male. 

```{r}
as.numeric(vec_sex)
```

This matters because when you plot them, factors will go with alphabetic order. 
## NSDUH
```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

table_marj = 
  read_html(nsduh_url) %>% 
  html_table() %>% 
  first() %>%
  slice(-1)
```

tidy up the NSDUH data
```{r}
data_marj = 
  table_marj %>%
  select(-contains("P Value")) %>%
  pivot_longer(
    -State, #this is to pivot anything except for "state"
    names_to = "age_year", 
    values_to = "percent") %>%
  separate(age_year, into = c("age", "year"), sep = "\\(") %>%
  mutate(
    year = str_replace(year, "\\)", ""), # get rid of the ")" in "year"
    percent = str_replace(percent, "[a-c]$", ""), # this is to get rid of the letters at the end in "percent". 
    percent = as.numeric(percent)) %>%
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West"))) #filter states that are not in these 5. 
```

Let’s quickly visualize these data for the 12-17 age group; to make the plot readable, we’ll treat State as a factor are reorder according to the median percent value.
```{r}
data_marj %>%
  filter(age == "12-17") %>% 
  mutate(State = fct_reorder(State, percent)) %>% #this is to recorder the x-axis states from alphabet order to order by percent. 
  ggplot(aes(x = State, y = percent, color = year)) + 
    geom_point() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


## Restaurant inspection
```{r}
data("rest_inspec")

rest_inspec %>% 
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  pivot_wider(names_from = grade, values_from = n)
```

```{r}
rest_inspec =
  rest_inspec %>%
  filter(grade %in% c("A", "B", "C"), boro != "Missing") %>% 
  mutate(boro = str_to_title(boro))
```
`str_to_title` we are leaving the first letter capitalized, and the other letters lower cases. 

Let's find pizza places
```{r}
rest_inspec %>% 
  filter(str_detect(dba, "Pizza")) %>% # this is to find pizza in dba column
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  pivot_wider(names_from = grade, values_from = n)

rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>%
  mutate(boro = fct_infreq(boro)) %>% #this is making the order of boro not by alphabetic order, so we convert to a factor label.
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar() 

rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>%
  mutate(
    boro = fct_infreq(boro),
    boro = str_replace(boro, "Manhattan", "The City")) %>% #Here, it can revert the boro back to string, so the sequence of these 2 lines matter. 
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar() 

rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>%
  mutate(
    boro = fct_infreq(boro),
    boro = replace(boro, which(boro == "Manhattan"), "The City")) %>% # or you can use replace. 
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()
```

