---
title: "lecture11.2_iteration & list cols"
author: "Fengdi Zhang"
date: "2022-11-03"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)

set.seed(1)
```

## lists

```{r}
vec_numeric = 5:8
vec_char = c("My", "name", "is", "Jeff")
vec_logical = c(TRUE, TRUE, TRUE, FALSE)
```

Let's put vectors in a list
```{r}
l = list(
  vec_numeric = 5:8,
  mat         = matrix(1:8, 2, 4), #matrix of 1 to 8, 2 rows, 4 cols. 
  vec_logical = c(TRUE, FALSE),
  summary     = summary(rnorm(1000)))
l
```

Accessing list items
```{r}
l$vec_numeric #using the dollar sign to pull out "vec_numeric"
l[[3]] #using double [[]] to pull out the third item in the list.
l[["mat"]] #using double [[]] to pull out the specific variable "mat"
```

## Loops!

Let's write a `for` loop

```{r}
list_norms = # first, create a list
  list(
    a = rnorm(20, 3, 1),
    b = rnorm(20, 0, 5),
    c = rnorm(20, 10, .2),
    d = rnorm(20, -3, 1)
  )

is.list(list_norms)
```

I want to calculate the mean and sd for each item in the list. 
```{r}
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Cannot be computed for length 1 vectors")
  }
  
  mean_x = mean(x)
  sd_x = sd(x)

  tibble(
    mean = mean_x, 
    sd = sd_x
  )
}

# we can do it separatly/repeatively. 
mean_and_sd(list_norms[[1]])
mean_and_sd(list_norms[[2]])
mean_and_sd(list_norms[[3]])
mean_and_sd(list_norms[[4]])
```

Let's use a `for` loop instead. 
```{r}
output = vector("list", length = 4)
# creating a list with 4 items, but the 4 spots are empty now. 

output[[1]] = mean_and_sd(list_norms[[1]])
#this is giving values for the first spot, but we want to do this for all 4 spots. 

for (i in 1:4) {
  output[[i]] = mean_and_sd(list_norms[[i]])
}

view(output)
```


## we can map

```{r}
map(list_norms, mean_and_sd)
map(list_norms, summary)
map(list_norms, median)
map(list_norms, var)

map_dbl(list_norms, median) # this returns a simplied output.
map_dfr(list_norms, mean_and_sd) # this is also a simplified output. collapse dataframes
```


## list columns..

```{r}
listcol_df = 
  tibble(
    name = c("a", "b", "c", "d"),
    samp = list_norms
  ) # each element tibble needs to have same length. 4 in this case. 

listcol_df[["samp"]] # this is pulling out the 4 normal distributions out. 

output = map(listcol_df[["samp"]], mean_and_sd)
```

```{r}
listcol_df %>% 
  mutate(summary = map(samp, mean_and_sd))
# within this column list, I want to calculate the mean and sd of each "samp", and save it as a new column "summary"
```


## What about something more realistic....Weather dataset

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```


let's nest within weather stations...
```{r}
weather_nest = 
  nest(weather_df, data = date:tmin)

weather_nest

weather_nest %>% pull(name)
weather_nest %>% pull(data)
weather_nest[["data"]]

```

```{r}
weather_nest[["data"]][[1]] # this is pulling out the first item from data from weather_nest
lm(tmax ~ tmin, data = weather_nest[["data"]][[1]])
#this is a regression for "tmax vs. tmin" using "weather_nest[["data"]][[1]]".
#Then we need to repeat it. 
lm(tmax ~ tmin, data = weather_nest[["data"]][[2]])
lm(tmax ~ tmin, data = weather_nest[["data"]][[3]])
```

we can write a function instead
```{r}
weather_lm = function(df) {
  lm(tmax ~ tmin, data = df)
}

weather_lm(weather_nest[["data"]][[1]])
```

Then put it in a map
```{r}
map(weather_nest[["data"]], weather_lm)
```

I want to add in a new column of the linear regression results. in a tidy way

```{r}
weather_nest = 
  weather_nest %>% 
  mutate(models = map(data, weather_lm))
```


## Unnesting
Now, I want to unnest the data column in the weather data. 
```{r}
weather_nest %>% 
  unnest(data)
```


## Napoleaon 
Here is my scraping function 
```{r}
read_page_reviews <- function(url) {
  
  html = read_html(url)
  
  review_titles = 
    html %>%
    html_nodes(".a-text-bold span") %>%
    html_text()
  
  review_stars = 
    html %>%
    html_nodes("#cm_cr-review_list .review-rating") %>%
    html_text() %>%
    str_extract("^\\d") %>%
    as.numeric()
  
  review_text = 
    html %>%
    html_nodes(".review-text-content span") %>%
    html_text() %>% 
    str_replace_all("\n", "") %>% 
    str_trim() %>% 
    str_subset("The media could not be loaded.", negate = TRUE) %>% 
    str_subset("^$", negate = TRUE)
  
  tibble(
    title = review_titles,
    stars = review_stars,
    text = review_text
  )
}
```

what we did last time
```{r}
url_base = "https://www.amazon.com/product-reviews/B00005JNBQ/ref=cm_cr_arp_d_viewopt_rvwer?ie=UTF8&reviewerType=avp_only_reviews&sortBy=recent&pageNumber="
vec_urls = str_c(url_base, 1:5)
```

```{r}
library(xml2)
map(vec_urls, read_page_reviews)
```


It can be even simpler
```{r}
dynamite_reviews = 
  tibble(
    page = 1:5,
    urls = str_c(url_base, page)) %>% 
  mutate(reviews = map(urls, read_page_reviews)) %>% 
  unnest()

```




