---
title: "9.27 Tidy Data"
author: "Fengdi Zhang"
date: "2022-09-27"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)

options(tibble.print_min = 5)
```

## R Markdown

load files
```{r}
pulse_df = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names()
```
we use the `haven` to read in SAS.Then, clean the names

## Make dataset tidy

```{r}
pulse_tidy_data = 
  pivot_longer(
    pulse_df, 
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi")
```
bdi_score_bl:bdi_score_12m: it's the columns you would like to pivot. 
names_to: the name we want in the newly combined column variable. 
values_to: the name we want in the newly combined column value.
names_prefix: the part that all names are same, so can be neglected. 

##Conduct the above steps in piping
```{r}
pulse_df = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  relocate(visit) %>%
  mutate(
    visit = replace(visit, visit == "bl", "00m"),
    visit = factor(visit)) %>%
  arrange(id, visit)
```
`mutate`: replace `visit=bl` to `visit=00m`.
`arrange(id, visit)`: first reorder by id, then by visit. 

##learning assessment
```{r}
litters_df = read_csv("data/FAS_litters.csv")%>%
  janitor::clean_names()%>%
  select(litter_number, ends_with("weight"))%>%
   pivot_longer(
    gd0_weight:gd18_weight,
    names_to = "gd", 
    values_to = "weight") %>% 
  mutate(gd = recode(gd, "gd0_weight" = 0, "gd18_weight" = 18))
```


##Convert long data into wide 
```{r}
analysis_result = tibble(
  group = c("treatment", "treatment", "placebo", "placebo"),
  time = c("pre", "post", "pre", "post"),
  mean = c(4, 8, 3.5, 4)
)

pivot_wider(
  analysis_result, 
  names_from = "time", 
  values_from = "mean")
```
names_from: which column you want to devide
values_from: the values you want to put.

##Binding Rows

First, import data from excel.
```{r}
fellowship_ring = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") %>%
  mutate(movie = "fellowship_ring") 
  #this is making a new row of "movie"

two_towers = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") %>%
  mutate(movie = "two_towers")

return_king = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") %>%
  mutate(movie = "return_king")
```

Use `bind_rows`
```{r}
lotr_tidy = 
  bind_rows(fellowship_ring, two_towers, return_king) %>%
  janitor::clean_names() %>%
  pivot_longer(
    female:male,
    names_to = "gender", 
    values_to = "words") %>%
  mutate(race = str_to_lower(race)) %>% #make "race" lower case#
  select(movie, everything()) #this is to reorder the columns.#
```


Joining datasets

read in 2 datasets first
```{r}
pup_data = 
  read_csv("./data/FAS_pups.csv") %>%
  janitor::clean_names() %>%
  mutate(
    sex = recode(sex, `1` = "male", `2` = "female"),
    sex = factor(sex)) 

litter_data = 
  read_csv("./data/FAS_litters.csv") %>%
  janitor::clean_names() %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3) %>%
  relocate(litter_number) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    dose = str_to_lower(dose))
```
`sep = 3` means we want to seperate after the third variable

Combine 2 datasets now
```{r}
fas_data = 
  left_join(pup_data, litter_data, by = "litter_number")
```
we do a left join because we want everything from the pup_data

rm(list=ls())

